---
title: "Code Along With Me (Episode 2)"
subtitle: "Performing basic exploratory data analysis with assistance from LLMs"
author: "William Okech"
date: "2024-06-20"
image: ""
#categories: [RStudio, R, Tutorial, Blog]
toc: true
format: 
  html:
    code-fold: show
    code-overflow: wrap
    warning: false
---

# Introduction

Using various LLMs for coding assistance

Prompt
Persona:  You are a data analyst working on exploratory data analysis at a newly funded startup. Context: You have been tasked with working on the penguins dataset and developing new techniques for EDA. Action: Write an R script to analyze the penguins the dataset, clean and process the data with tidyverse, perform EDA with plots. Tone: Professional. Output Format: Provide an R script.

EDA is very subjective so outputs will differ

Here we tested whether the code could be run as-is and the number of modifications required. 

# 1. ChatGPT

## Corrections to code:
a. Needed to install GGally before the code worked

## Code

```{r, results='hide'}

# Load necessary libraries
library(tidyverse)
library(palmerpenguins)
library(ggplot2)
library(GGally) # Was not included in the original code output

# Load the penguins dataset
data("penguins")

# Preview the dataset
glimpse(penguins)

# Data Cleaning and Processing
# Remove rows with missing values
penguins_clean <- penguins %>%
  drop_na()

# Convert necessary columns to appropriate data types
penguins_clean <- penguins_clean %>%
  mutate(species = factor(species),
         island = factor(island),
         sex = factor(sex))

# Summary statistics
summary(penguins_clean)
```

```{r}
# Exploratory Data Analysis (EDA)

# Distribution of species
ggplot(penguins_clean, aes(x = species)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Distribution of Penguin Species", x = "Species", y = "Count")

# Distribution of flipper length by species
ggplot(penguins_clean, aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_boxplot() +
  labs(title = "Flipper Length by Species", x = "Species", y = "Flipper Length (mm)") +
  theme(legend.position = "none")

# Relationship between flipper length and body mass
ggplot(penguins_clean, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point(alpha = 0.7) +
  labs(title = "Relationship between Flipper Length and Body Mass", 
       x = "Flipper Length (mm)", y = "Body Mass (g)")

# Pairwise scatter plots
penguins_clean %>%
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g) %>%
  ggpairs(aes(color = penguins_clean$species, alpha = 0.7),
          upper = list(continuous = wrap("cor", size = 3)),
          lower = list(continuous = wrap("points", size = 0.5))) +
  theme_bw() +
  labs(title = "Pairwise Scatter Plots of Penguin Measurements")

# Density plot of body mass by species and sex
ggplot(penguins_clean, aes(x = body_mass_g, fill = species)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ sex) +
  labs(title = "Density Plot of Body Mass by Species and Sex", 
       x = "Body Mass (g)", y = "Density")

# Save cleaned dataset for further analysis
# write_csv(penguins_clean, "cleaned_penguins_dataset.csv")

# Additional analysis can be performed as needed

```

# 2. CoPilot

## Corrections to code:
a. library(palmerpenguins) was not included, instead this line of code was provided: penguins <- read.csv("penguins.csv")
b. 

```{r, results='hide'}

# Load necessary libraries
library(tidyverse)
library(palmerpenguins)

# Read the penguins dataset (assuming it's already loaded or available)
# Replace 'penguins.csv' with the actual file path if needed
# Load the penguins dataset

data("penguins")

# Check the structure of the dataset
str(penguins)

# Summary statistics
summary(penguins)

# Data cleaning: Remove missing values
penguins_clean <- penguins %>%
  drop_na()

```

```{r}
# EDA: Create scatter plots
ggplot(penguins_clean, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point() +
  labs(x = "Bill Length (mm)", y = "Bill Depth (mm)", title = "Penguins Bill Measurements")

# Boxplots
ggplot(penguins_clean, aes(x = species, y = flipper_length_mm)) +
  geom_boxplot() +
  labs(x = "Species", y = "Flipper Length (mm)", title = "Flipper Length by Species")

# Histograms
ggplot(penguins_clean, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 100, fill = "skyblue", color = "black") +
  labs(x = "Body Mass (g)", y = "Frequency", title = "Distribution of Body Mass")

# Correlation matrix
cor_matrix <- cor(penguins_clean[, c("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g")])
print(cor_matrix)

# Pairwise scatter plots
pairs(penguins_clean[, c("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g")])

# Additional EDA steps can be added as needed

# Save plots (optional)
#ggsave("penguins_scatter.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
#ggsave("penguins_boxplot.png", plot = last_plot(), width = 8, height = 6, dpi = 300)
#ggsave("penguins_histogram.png", plot = last_plot(), width = 8, height = 6, dpi = 300)

# End of script
```

# 3. Google Gemini

## Corrections to code:
1. Needed a few more iterations for the code to work without errors
2. 

```{r, results='hide'}

# Load libraries
library(palmerpenguins)
library(tidyverse)

# Load penguin data
penguins <- palmerpenguins::penguins

# Check for missing values
summary(penguins)

# Handle missing values (consider imputation or removal based on analysis needs)
# This example removes rows with missing values
penguins <- na.omit(penguins)

# Explore data distribution
penguins %>%
  group_by(species) %>%
  summarise(mean_bill_length = mean(bill_length_mm),
             sd_bill_length = sd(bill_length_mm))

```

```{r}
# Visualize distribution of bill length by species
penguins %>%
  ggplot(aes(x = bill_length_mm, color = species)) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution of Bill Length by Species",
       x = "Bill Length (mm)",
       y = "Density") +
  facet_wrap(~species)

# Explore relationships between numerical variables
penguins %>%
  ggplot(aes(x = bill_depth_mm, y = flipper_length_mm, color = species)) +
  geom_point(alpha = 0.5) +
  labs(title = "Bill Depth vs. Flipper Length by Species",
       x = "Bill Depth (mm)",
       y = "Flipper Length (mm)") +
  facet_wrap(~species)

# Explore categorical relationships
penguins %>%
  ggplot(aes(x = island, fill = species)) +
  geom_bar(stat = "count") +
  labs(title = "Number of Penguins by Island and Species",
       x = "Island",
       y = "Count")

# Explore body size differences by species
penguins %>%
  group_by(species) %>%
  summarise(mean_body_mass = mean(body_mass_g)) %>%
  ggplot(aes(x = reorder(species, mean_body_mass), y = mean_body_mass)) +
  geom_col(fill = "skyblue") +
  labs(title = "Average Body Mass by Species",
       x = "Species",
       y = "Body Mass (g)")

# Further exploration (consider additional plots or tests)
# * Correlation analysis between numerical variables
# * Boxplots to compare distributions
# * Hypothesis testing for group differences

# Remember to adapt and expand this script based on your specific research questions
```


# 4. Anthropic Claude

## Corrections to code:
1. Worked well
2. 

```{r, results='hide'}

# Load required packages
library(tidyverse)
library(palmerpenguins)

# Load the penguins dataset
data <- penguins

# View the data structure
str(data)

# Check for missing values
sum(is.na(data))

# Drop rows with missing values
data <- na.omit(data)

# Data cleaning and preprocessing
data <- data %>%
 mutate(species = as.factor(species),
        island = as.factor(island),
        sex = as.factor(sex))

# Exploratory Data Analysis

# Summary statistics
summary(data)

```

```{r}

# Visualize species distribution
ggplot(data, aes(x = species)) +
 geom_bar() +
 labs(title = "Species Distribution",
      x = "Species",
      y = "Count")

# Visualize bill length vs bill depth
ggplot(data, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
 geom_point() +
 labs(title = "Bill Length vs Bill Depth",
      x = "Bill Length (mm)",
      y = "Bill Depth (mm)",
      color = "Species")

# Visualize body mass distribution
ggplot(data, aes(x = body_mass_g, fill = species)) +
 geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
 labs(title = "Body Mass Distribution",
      x = "Body Mass (g)",
      y = "Count",
      fill = "Species")

# Visualize flipper length by species and sex
ggplot(data, aes(x = species, y = flipper_length_mm, fill = sex)) +
 geom_boxplot() +
 labs(title = "Flipper Length by Species and Sex",
      x = "Species",
      y = "Flipper Length (mm)",
      fill = "Sex")

```

# 5. Perplexity

## Corrections to code:
1. Typo: "flippers_length_mm"
2. Initially didn't color the plots

```{r, results='hide'}

# Load necessary libraries
library(tidyverse)
library(corrplot)

# Load the penguins dataset
data("penguins", package = "palmerpenguins")

# View the first few rows of the dataset
head(penguins)

# Clean and process the data
penguins <- penguins %>%
  mutate(
    species = factor(species),
    sex = factor(sex),
    island = factor(island)
  )

# Perform exploratory data analysis
# Summary statistics
summary_stats <- penguins %>%
  group_by(species) %>%
  summarise(
    mean_bill_length = mean(bill_length_mm),
    mean_bill_depth = mean(bill_depth_mm),
    mean_flipper_length = mean(flipper_length_mm),
    mean_body_mass = mean(body_mass_g)
  )
```

```{r}

# Visualize the data
#   a. Distribution of bill length
ggplot(penguins, aes(x = bill_length_mm)) +
  geom_histogram(binwidth = 1, color = "black", fill = "lightblue") +
  labs(title = "Distribution of Bill Length", x = "Bill Length (mm)", y = "Frequency")

#   b. Distribution of bill depth
ggplot(penguins, aes(x = bill_depth_mm)) +
  geom_histogram(binwidth = 1, color = "black", fill = "lightblue") +
  labs(title = "Distribution of Bill Depth", x = "Bill Depth (mm)", y = "Frequency")

#   c. Distribution of flipper length
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram(binwidth = 1, color = "black", fill = "lightblue") +
  labs(title = "Distribution of Flipper Length", x = "Flipper Length (mm)", y = "Frequency")

#   d. Distribution of body mass
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 100, color = "black", fill = "lightblue") +
  labs(title = "Distribution of Body Mass", x = "Body Mass (g)", y = "Frequency")

#   e. Species distribution
ggplot(penguins, aes(x = species, fill = species)) +
  geom_bar() +
  labs(title = "Species Distribution", x = "Species", y = "Count", fill = "Species")

#   f. Sex distribution
ggplot(penguins, aes(x = sex, fill = sex)) +
  geom_bar() +
  labs(title = "Sex Distribution", x = "Sex", y = "Count", fill = "Sex")

#   g. Island distribution
ggplot(penguins, aes(x = island, fill = island)) +
  geom_bar() +
  labs(title = "Island Distribution", x = "Island", y = "Count", fill = "Island")

```

# 6. Cohere

## Corrections to code:
1. Loaded data with penguins <- read_csv("penguins.csv")
2. Typo: gguins_cleaned
3. Doesn't segment species in scatter plots by color

```{r, results='hide'}

# Load the required packages
library(tidyverse)
library(ggplot2)
library(readr)

# Read the penguins dataset

data("penguins", package = "palmerpenguins")

# Explore the structure of the dataset
str(penguins)

# Summary statistics for numerical variables
summary(penguins[sapply(penguins, is.numeric)])

# Check for missing values
is.na(penguins)

# Data Cleaning and Processing

# Convert bill_length_mm and bill_depth_mm to numeric
penguins <- penguins %>%
  mutate(across(c(bill_length_mm, bill_depth_mm), as.numeric))

# Handle missing values
penguins_cleaned <- penguins %>%
  drop_na()

# Summary statistics after cleaning
summary(penguins_cleaned[sapply(penguins_cleaned, is.numeric)])

```

```{r}

# Exploratory Data Analysis with Plots

# Basic scatter plot of bill length vs bill depth
basic_scatter <- ggplot(penguins_cleaned, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  ggtitle("Bill Length vs Bill Depth")

print(basic_scatter)

# Box plot of bill depth by species
box_plot <- ggplot(penguins_cleaned, aes(x = species, y = bill_depth_mm)) +
  geom_boxplot() +
  ggtitle("Bill Depth by Species")

print(box_plot)

# Histogram of flipper length
histogram <- ggplot(penguins_cleaned, aes(x = flipper_length_mm)) +
  geom_histogram(binwidth = 5) +
  ggtitle("Distribution of Flipper Length")

print(histogram)

# Density plot of body mass by species
density_plot <- ggplot(penguins_cleaned, aes(x = body_mass_g, fill = species)) +
  geom_density(alpha = 0.7) +
  ggtitle("Density Plot of Body Mass by Species")

print(density_plot)

```